---
title: "HDHP model codeshare"
author: "Sanjay Basu"
date: "1/21/2019"
output: html_document
---

```{r setup, include=FALSE}
#rm(list=ls())
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, warning = FALSE, message = FALSE, cache.lazy = FALSE)
library(foreign)
library(survey)
library(tidyverse)
library(plyr)
library(ggplot2)
library(scales)
library(nhanesA)
library(stringr)


```

## KFF 2018 EHBS

```{r import}
df = read.spss('~/Box/Analytics Team/Data Sources & Examples/KFF Health Benefits Survey/health benefits 2018/Health Benefits 2018.sav', to.data.frame = T)

```

## descriptive stats

```{r desc}
options( survey.lonely.psu = "adjust" )

dftemp = data.frame(enroll= c(df$b12c,df$b12e),
                    type = c(rep("PPO",length(df$b12c)), rep("HDHP",length(df$b12e))),
                    empwt = df$empwt)
mu = data.frame(grp.mean = c(.49,.29),
                type = c("PPO","HDHP"))
p<-ggplot(dftemp, aes(x=enroll, color=type, weight = empwt)) +
  geom_histogram(fill="white", position="dodge")+
  geom_vline(data=mu, aes(xintercept=grp.mean, color=type),
             linetype="dashed")
p + scale_color_brewer(palette="Set1") + 
  theme_minimal()+theme(legend.position="top") +  scale_y_continuous(labels = scales::comma)+
  labs(title="Proportion of workers subscribing to HDHP or PPO plan",x="Proportion of workers enrolled in each firm", y = "Number of firms")


dftemp = data.frame(enroll= c(df$ddctsppo,df$ddctshdp),
                    type = c(rep("PPO",length(df$ddctsppo)), rep("HDHP",length(df$ddctshdp))),
                    empwt = df$empwt)
mu = data.frame(grp.mean = c(1204,2349),
                type = c("PPO","HDHP"))
p<-ggplot(dftemp, aes(x=enroll, color=type, weight = empwt)) +
  geom_histogram(fill="white", position="dodge")+
  geom_vline(data=mu, aes(xintercept=grp.mean, color=type),
             linetype="dashed")
p + scale_color_brewer(palette="Set1") + 
  theme_minimal()+theme(legend.position="top") +  scale_y_continuous(labels = scales::comma)+
  labs(title="Deductibles for Single Coverage",x="$US", y = "Number of firms")


dftemp = data.frame(enroll= c(df$prmsppoa,df$prmshdpa),
                    type = c(rep("PPO",length(df$prmsppoa)), rep("HDHP",length(df$prmshdpa))),
                    empwt = df$empwt)
mu = data.frame(grp.mean = c(7149,6459),
                type = c("PPO","HDHP"))
p<-ggplot(dftemp, aes(x=enroll, color=type, weight = empwt)) +
  geom_histogram(fill="white", position="dodge")+
  geom_vline(data=mu, aes(xintercept=grp.mean, color=type),
             linetype="dashed")
p + scale_color_brewer(palette="Set1") + 
  theme_minimal()+theme(legend.position="top") +  scale_y_continuous(labels = scales::comma)+
  labs(title="Premiums for Single Coverage",x="$US", y = "Number of firms")



dftemp = data.frame(enroll= c(df$g43ann),
                    type = c(rep("HDHP",length(df$g43ann))),
                    hsawt = df$hsawt)
mu = data.frame(grp.mean = c(603),
                type = c("HDHP"))
p<-ggplot(dftemp, aes(x=enroll, color=type, weight = hsawt)) +
  geom_histogram(fill="white", position="dodge")+
  geom_vline(data=mu, aes(xintercept=grp.mean, color=type),
             linetype="dashed")
p + scale_color_brewer(palette="Set1") + 
  theme_minimal()+theme(legend.position="top") +  scale_y_continuous(labels = scales::comma)+
  labs(title="HSA Firm Contribution",x="$US", y = "Number of firms")


```




## employer costs

premium change: calculated from Lo Sasso table 4 column 2 [OLS] how much change in total spending would occur without HSA and with lower deductible including in-network coinsurance, and hosp/surgery deductible (significant variables at 5% alpha), with an 85% loss ratio from claims to premium
DOI: 10.1111/j.1539-6975.2009.01346.x


```{r emp cost}

covwt.design <- svydesign(
      ~1 ,
      strata = ~newcell ,
      weight = ~hsawt ,
      data = subset( df , hsawt > 0 )
      )

# Annual HSA Firm Contribution: Single - HDP
svymean(~g43ann, covwt.design,na.rm=T)[1]
svyquantile(~g43ann, covwt.design,na.rm=T, quantiles = c(.025,.5,.975))

covwt.design <- svydesign(
      ~1 ,
      strata = ~newcell ,
      weight = ~hdpwt ,
      data = subset( df , hdpwt > 0 )
      )

# Average Deductible Among Covered Workers Who Face a Deductible for Single Coverage - HDP
svymean(~ddctshdp, covwt.design,na.rm=T)[1]
svyquantile(~ddctshdp, covwt.design,na.rm=T, quantiles = c(.025,.5,.975))


covwt.design <- svydesign(
      ~1 ,
      strata = ~newcell ,
      weight = ~ppowt ,
      data = subset( df , ppowt > 0 )
      )

# Average Deductible Among Covered Workers Who Face a Deductible for Single Coverage - PPO
svymean(~ddctsppo, covwt.design,na.rm=T)[1]
svyquantile(~ddctsppo, covwt.design,na.rm=T, quantiles = c(.025,.5,.975))


# Within-firm difference in deductible
summary(df$ddctshdp-df$ddctsppo)


covwt.design <- svydesign(
      ~1 ,
      strata = ~newcell ,
      weight = ~hdpwt ,
      data = subset( df , hdpwt > 0 )
      )


#Annual Premiums for Single coverage - HDP
svymean(~prmshdpa, covwt.design,na.rm=T)[1] 
svyquantile(~prmshdpa, covwt.design,na.rm=T, quantiles = c(.025,.5,.975))


covwt.design <- svydesign(
      ~1 ,
      strata = ~newcell ,
      weight = ~ppowt ,
      data = subset( df , ppowt > 0 )
      )

svymean(~prmsppoa, covwt.design,na.rm=T)[1] 
svyquantile(~prmsppoa, covwt.design,na.rm=T, quantiles = c(.025,.5,.975))


covwt.design <- svydesign(
      ~1 ,
      strata = ~newcell ,
      weight = ~hdpwt ,
      data = subset( df , hdpwt > 0 )
      )

# Annual Contribution for Single coverage - HDP
svymean(~wkrshdpa, covwt.design,na.rm=T)[1]
svyquantile(~wkrshdpa, covwt.design,na.rm=T, quantiles = c(.025,.5,.975))

covwt.design <- svydesign(
      ~1 ,
      strata = ~newcell ,
      weight = ~ppowt ,
      data = subset( df , ppowt > 0 )
      )
# Annual Contribution for Single coverage - PPO
svymean(~wkrsppoa, covwt.design,na.rm=T)[1]
svyquantile(~wkrsppoa, covwt.design,na.rm=T, quantiles = c(.025,.5,.975))


```



# health impact
then calculate impact of reduced adherence to DM, statins, and HTN drugs based on 5% improvement in adherence [ https://doi.org/10.1161/CIRCOUTCOMES.118.004632 ]
given baseline rates of: 68.6% for diabetes medicines, 71.8% for statins, and 71.6% for blood pressure drugs [ https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2842539/ , table 4 column 5 ]

and given meta-analytic estimates of delta A1c => RECODE estimates of change in complications and associated $ over time until retirement, and calculation of decr MI/stroke risk and Dollars associated w statin or HTN drugs given AHA/ACC ASCVD Risk estimator + online and NHANES employed population as simualted cohort, subset to the NHANES subpop that has the prevalence of HDHP enrollees per NHIS


```{r nhanes nhis}

# NHANES 2015-2016


demo <- nhanes('DEMO_I')
bmx <- nhanes('BMX_I')
tchol <- nhanes('TCHOL_I')
hdl <- nhanes('HDL_I')
bpq <- nhanes('BPQ_I')
ghb <- nhanes('GHB_I')
diq <- nhanes('DIQ_I')
biopro <- nhanes('BIOPRO_I')
albcr <- nhanes('ALB_CR_I')
bpx <- nhanes('BPX_I')
diet <- nhanes('DR1IFF_I')
rxq <- nhanes('RXQ_RX_I')
smq <- nhanes('SMQ_I')
mcq <- nhanes('MCQ_I')
occ <- nhanes('OCQ_I')
glu <- nhanes('GLU_I')
hic <- nhanes('HIQ_I')

demo <- nhanesTranslate('DEMO_I', 
                         colnames = c('RIAGENDR','RIDRETH1'), 
                         data=demo)



merged = merge(demo, bmx)
merged = merge(merged, tchol)
merged = merge(merged, hdl)
merged = merge(merged, bpq)
merged = merge(merged, ghb)
merged = merge(merged, diq)
merged = merge(merged, biopro)
merged = merge(merged, albcr)
merged = merge(merged, bpx)
merged = merge(merged, diet) 
merged = merge(merged, rxq)
merged = merge(merged, smq)
merged = merge(merged, mcq)
merged = merge(merged, occ)
merged = merge(merged, glu)
merged = merge(merged, hic)
mergedsub = merged %>%
  dplyr::filter(HIQ031A==14) # privately-insured subset


age = mergedsub$RIDAGEYR
female = (mergedsub$RIAGENDR=="Female")
black = (mergedsub$RIDRETH1=="Non-Hispanic Black")
black[is.na(black)]=0
hisp = (mergedsub$RIDRETH1=="Other Hispanic")
hisp[is.na(hisp)]=0
bmi = mergedsub$BMXBMI
bmi[is.na(bmi)]=mean(na.omit(bmi))
totchol = mergedsub$LBXTC
totchol[is.na(totchol)]=mean(na.omit(totchol))
hdlc = mergedsub$LBDHDD
hdlc[is.na(hdlc)]=mean(na.omit(hdlc))
dm = (mergedsub$DIQ010==1) | 
  (mergedsub$LBXGH>=6.5) | 
  (mergedsub$LBXGLU>=126)
dm[is.na(dm)]=0
sercreat = mergedsub$LBXSCR
sercreat[is.na(sercreat)]=mean(na.omit(sercreat))
uralbcreat =  mergedsub$URDACT
uralbcreat[is.na(uralbcreat)]=mean(na.omit(uralbcreat))
sysbp = mergedsub$BPXSY1
sysbp[is.na(sysbp)]=mean(na.omit(sysbp))
rxbp = (mergedsub$BPQ040A==1)
rxbp[is.na(rxbp)]=0
cvdhist = (mergedsub$MCQ160F==1)|(mergedsub$MCQ160F==1)
cvdhist[is.na(cvdhist)]=0
hgba1c = mergedsub$LBXGH
hgba1c[is.na(hgba1c)]=mean(na.omit(hgba1c))
cursmoke = (mergedsub$SMQ040==1 | mergedsub$SMQ040==2)
cursmoke[is.na(cursmoke)]=0

statin = grepl("STATIN",mergedsub$RXDDRUG)
statin[is.na(statin)]=0
oralrx = grepl("METFORMIN",mergedsub$RXDDRUG)|grepl("GLIPIZIDE",mergedsub$RXDDRUG)
oralrx[is.na(oralrx)]=0
anticoag = grepl("WARFARIN",mergedsub$RXDDRUG)
anticoag[is.na(anticoag)]=0



```

# Calculate baseline risk scores, 10 year => project to pre-retirement

1. Yadlowsky S, Hayward RA, Sussman JB, McClelland RL, Min Y, Basu S. Clinical Implications of Revised Pooled Cohort Equations for Estimating Atherosclerotic Cardiovascular Disease Risk. Ann Intern Med. ;169:20–29. doi: 10.7326/M17-3011

2. Basu S, Sussman JB, Berkowitz SA, Hayward RA, Yudkin JS. Development and validation of Risk Equations for Complications Of type 2 Diabetes (RECODe) using individual participant data from randomised trials. Lancet Diabetes Endocrinol. 2017;5(10):788-798.

3. Sanjay Basu, Jeremy B. Sussman, Seth A. Berkowitz, Rodney A. Hayward, Alain G. Bertoni, Adolfo Correa, Stanford Mwasongwe, John S. Yudkin. Validation of Risk Equations for Complications of Type 2 Diabetes (RECODe) Using Individual Participant Data From Diverse Longitudinal Cohorts in the U.S. Diabetes Care Mar 2018, 41 (3) 586-595; DOI: 10.2337/dc17-2002




```{r baseline risk scores, 10 year}

# ASCVD
ascvd_10   = ( 1.0*female / (1.0 + exp( - (
  -12.823110 +
    0.106501 * as.numeric(age) +
    0.432440 * as.numeric(black) +
    0.000056 * (as.numeric(sysbp) ^ 2) +
    0.017666 * as.numeric(sysbp) +
    0.731678 * as.numeric(rxbp) +
    0.943970 * as.numeric(dm) +
    1.009790 * as.numeric(cursmoke) +
    0.151318 * (as.numeric(totchol) / as.numeric(hdlc)) +
    -0.008580 * as.numeric(age) * as.numeric(black) +
    -0.003647 * as.numeric(sysbp) * as.numeric(rxbp) +
    0.006208 * as.numeric(sysbp) * as.numeric(black) +
    0.152968 * as.numeric(black) * as.numeric(rxbp) +
    -0.000153 * as.numeric(age) * as.numeric(sysbp) +
    0.115232 * as.numeric(black) * as.numeric(dm) +
    -0.092231 * as.numeric(black) * as.numeric(cursmoke) +
    0.070498 * as.numeric(black) * (as.numeric(totchol) / as.numeric(hdlc)) +
    -0.000173 * as.numeric(black)  * as.numeric(sysbp) * as.numeric(rxbp) +
    -0.000094 * as.numeric(age) * as.numeric(sysbp) * as.numeric(black)))))+
  (1.0*(1-female) / (1.0 + exp( - (
  -11.679980 +
    0.064200 * as.numeric(age) +
    0.482835 * as.numeric(black) +
    -0.000061 * (as.numeric(sysbp) ^ 2) +
    0.038950 * as.numeric(sysbp) +
    2.055533 * as.numeric(rxbp) +
    0.842209 * as.numeric(dm) +
    0.895589 * as.numeric(cursmoke) +
    0.193307 * (as.numeric(totchol) / as.numeric(hdlc)) +
    -0.014207 * as.numeric(sysbp) * as.numeric(rxbp) +
    0.011609 * as.numeric(sysbp) * as.numeric(black) +
    -0.119460 * as.numeric(rxbp) * as.numeric(black) +
    0.000025 * as.numeric(age) * as.numeric(sysbp) +
    -0.077214 * as.numeric(black) * as.numeric(dm) +
    -0.226771 * as.numeric(black) * as.numeric(cursmoke) +
    -0.117749 * (as.numeric(totchol) / as.numeric(hdlc)) * as.numeric(black) +
    0.004190 * as.numeric(black) * as.numeric(rxbp) * as.numeric(sysbp) +
    -0.000199 * as.numeric(black) * as.numeric(age) * as.numeric(sysbp)))))

      
# DM incidence
# https://www.cdc.gov/diabetes/data/statistics-report/incidence-diabetes.html
dminc_10 = (mean(c(.031,.068))*(age<45)+
  mean(c(.109,.068))*(age>=45 & age<65)+
  mean(c(.094,.068))*(age>=65))*100/1000*10


# DM neph
nbetax = (-1.938e-02*age+
           -1.129e-02*female+
           -8.812e-02*black+
           2.338e-01*hisp+
           1.483e-01*cursmoke+
           3.027e-03*sysbp+
           -7.952e-02*rxbp+
           -1.256e-01*oralrx+
           3.199e-02*anticoag+
           -2.164e-02*cvdhist+
           1.369e-01*hgba1c+
           -1.112e-03*totchol+
           6.289e-03*hdlc+
           8.609e-01*sercreat+
           3.615e-04*uralbcreat)
esrdrisk_10 = 1-.973^exp(nbetax-mean(nbetax,na.rm=T))

# DM retin

rbetax = (2.285e-02*age+
           2.264e-01*female+
           -1.677e-01*black+
           8.243e-03*sysbp+
           6.393e-02*rxbp+
           -2.349e-01*oralrx+
           1.127e-01*cvdhist+
           1.449e-01*hgba1c+
           -1.676e-04*totchol+
           5.447e-03*hdlc+
           6.947e-01*sercreat+
           1.992e-04*uralbcreat)
retinrisk_10 = 1 - .92^exp(rbetax-mean(rbetax,na.rm=T))


# DM neuro
ebetax = (0.0302237*age+
           -0.1868000*female+
            -0.0944841*black+
           0.0045609*sysbp+
           0.1819157*rxbp+
           -0.2574724*oralrx+
           0.2667152*cvdhist+
           0.1886579*hgba1c+
           0.0021850*totchol+
           -0.0053887*hdlc+
           0.6044183*sercreat)
neurorisk_10 = 1 - .87^exp(ebetax-mean(ebetax,na.rm=T))



# All cause mort
abetax = (6.703e-02*age+
           -1.529e-01*female+
            -2.393e-02*black+
           5.399e-01*cursmoke+
           -2.988e-03*sysbp+
           8.776e-02*rxbp+
           -2.681e-01*statin+
           4.036e-01*anticoag+
           -0.2574724*oralrx+
           5.888e-01*cvdhist+
           1.659e-01*hgba1c+
           -9.478e-04*totchol+
           -4.378e-03*hdlc+
           3.597e-01*sercreat+
           3.889e-04*uralbcreat)
mortrisk_10 = 1 - .93^exp(abetax-mean(abetax,na.rm=T))

# breast Ca,  https://gis.cdc.gov/Cancer/USCS/DataViz.html
bcarisk_10 = ((125/100000)*(1-hisp)*(1-black)+
  (93.7/100000)*hisp+
  (122.6/100000)*black+
  (0.2/100000)*(age<20)+
  (1.7/100000)*(age>=20 & age<25)+
  (10.3/100000)*(age>=25 & age<30)+
  (30.3/100000)*(age>=30 & age<35)+
  (62.2/100000)*(age>=35 & age<40)+
  (122.6/100000)*(age>=40 & age<45)+
  (188.4/100000)*(age>=45 & age<50)+
  (226.6/100000)*(age>=50 & age<55)+
  (263.3/100000)*(age>=55 & age<60)+
  (328.6/100000)*(age>=60 & age<65)+
  (418.6/100000)*(age>=65 & age<70)+
  (460.1/100000)*(age>=70 & age<75)+
  (445.1/100000)*(age>=75))/2

  
```

```{r baseline risk scores, through pre-retirement}
# https://www.cdc.gov/nchs/data/nvsr/nvsr66/nvsr66_04.pdf
lifeexp = (female==0)*(black==0)*(hisp==0)*( # NH WM (non-Hispanic White Male)
62.1*(age<20)+
57.3*(age>=20 & age<25)+
52.6*(age>=25 & age<30)+
47.9*(age>=30 & age<35)+
43.3*(age>=35 & age<40)+
38.7*(age>=40 & age<45)+
34.2*(age>=45 & age<50)+
29.8*(age>=50 & age<55)+
25.7*(age>=55 & age<60)+
21.7*(age>=60 & age<65)+
18*(age>=65 & age<70)+
14.4*(age>=70 & age<75)+
11.2*(age>=75 & age<80)+
8.3*(age>=80)
)+
  (female==1)*(black==0)*(hisp==0)*( # NH WF
66.7*(age<20)+
61.8*(age>=20 & age<25)+
56.9*(age>=25 & age<30)+
52.1*(age>=30 & age<35)+
47.3*(age>=35 & age<40)+
42.5*(age>=40 & age<45)+
37.9*(age>=45 & age<50)+
33.3*(age>=50 & age<55)+
28.9*(age>=55 & age<60)+
24.7*(age>=60 & age<65)+
20.5*(age>=65 & age<70)+
16.6*(age>=70 & age<75)+
13*(age>=75 & age<80)+
9.7*(age>=80)
  )+
  (female==0)*(black==1)*(hisp==0)*( # NH BM
58.3*(age<20)+
53.6*(age>=20 & age<25)+
49*(age>=25 & age<30)+
44.5*(age>=30 & age<35)+
40*(age>=35 & age<40)+
35.6*(age>=40 & age<45)+
31.2*(age>=45 & age<50)+
27*(age>=50 & age<55)+
23*(age>=55 & age<60)+
19.4*(age>=60 & age<65)+
16.3*(age>=65 & age<70)+
13.2*(age>=70 & age<75)+
10.5*(age>=75 & age<80)+
8.1*(age>=80)  
)+
  (female==1)*(black==1)*(hisp==0)*( # NH BF
64.2*(age<20)+
59.3*(age>=20 & age<25)+
54.5*(age>=25 & age<30)+
49.7*(age>=30 & age<35)+
44.9*(age>=35 & age<40)+
40.3*(age>=40 & age<45)+
35.7*(age>=45 & age<50)+
31.3*(age>=50 & age<55)+
27.2*(age>=55 & age<60)+
23.3*(age>=60 & age<65)+
19.5*(age>=65 & age<70)+
16*(age>=70 & age<75)+
12.7*(age>=75 & age<80)+
9.8*(age>=80)      
  )+
  (female==0)*(black==0)*(hisp==1)*( # HM
65*(age<20)+
60.1*(age>=20 & age<25)+
55.4*(age>=25 & age<30)+
50.7*(age>=30 & age<35)+
45.9*(age>=35 & age<40)+
41.2*(age>=40 & age<45)+
36.6*(age>=45 & age<50)+
32.1*(age>=50 & age<55)+
27.7*(age>=55 & age<60)+
23.6*(age>=60 & age<65)+
19.7*(age>=65 & age<70)+
16*(age>=70 & age<75)+
12.6*(age>=75 & age<80)+
9.5*(age>=80)    
)+
  (female==1)*(black==0)*(hisp==1)*( # HF
70*(age<20)+
65*(age>=20 & age<25)+
60.2*(age>=25 & age<30)+
55.3*(age>=30 & age<35)+
50.4*(age>=35 & age<40)+
45.5*(age>=40 & age<45)+
40.7*(age>=45 & age<50)+
36.1*(age>=50 & age<55)+
31.5*(age>=55 & age<60)+
27*(age>=60 & age<65)+
22.8*(age>=65 & age<70)+
18.7*(age>=70 & age<75)+
14.8*(age>=75 & age<80)+
11.2*(age>=80)       
  )

lifeexp_prop = (78.69-65)/lifeexp
lifeexp_prop[lifeexp_prop>(1-1/min(lifeexp))] = (1-1/min(lifeexp))
lifeexp[lifeexp<0] = 0

ascvd_rate = -log(1-ascvd_10)/10
ascvd_life = 1-exp(-ascvd_rate*lifeexp)

dminc_rate = -log(1-dminc_10)/10
dminc_life = 1-exp(-dminc_rate*lifeexp)

esrdrisk_rate = -log(1-esrdrisk_10)/10
esrdrisk_life = 1-exp(-esrdrisk_rate*lifeexp)

retinrisk_rate = -log(1-retinrisk_10)/10
retinrisk_life = 1-exp(-retinrisk_rate*lifeexp)

neurorisk_rate = -log(1-neurorisk_10)/10
neurorisk_life = 1-exp(-neurorisk_rate*lifeexp)

mortrisk_rate = -log(1-mortrisk_10)/10
mortrisk_life = 1-exp(-mortrisk_rate*lifeexp)

bca_rate = -log(1-bcarisk_10)/10
bca_life = 1-exp(-bca_rate*lifeexp)

```

# after 5% improvement in adherence
supplemental table 3 from: https://doi.org/10.1177/2381468317725741

BP meds:
2^((-8.8*((-1.84775*(10^(-5)))*(57.88^2)+(1.5841*(10^(-3)))*57.88+(2.8672*(10^(-2))))))
 = 0.70
RR reduced by 0.70 for 5% 

statins:
RR reduced by 30% [atorva20] for 5% 

diabetes meds:
appendix page 4 from: https://dx.doi.org/10.1016%2FS2213-8587(16)30270-4 
0.77 for retinal photocoagulation (95% CI: 0.61, 0.97), 
0.64 for amputation of lower extremity (95% CI: 0.43, 0.95), 
0.78 for nephropathy (95% CI: 0.61, 0.99) for a 0.9% reduction in A1c.
average A1c reduction per med = 1.5%
for 5% 

mammography screening:
doi: 10.1200/JCO.2017.75.2501
HDHP members have delayed early-stage breast cancer diagnosis (aHR, 0.83; 0.78 to 0.90)

table 1 from: https://annals.org/aim/fullarticle/2480757/screening-breast-cancer-u-s-preventive-services-task-force-



```{r postadh}

ascvd_life_post = ascvd_life*(1-0.7*0.7*.05)

esrdrisk_life_post = esrdrisk_life*(1-0.78*.05)

retinrisk_life_post = retinrisk_life*(1-0.77*.05)

neurorisk_life_post = neurorisk_life*(1-0.64*.05)

bca_life_post = bca_life*(0.83)


summary(neurorisk_life_post)
quantile(neurorisk_life_post,c(.025,.975))
summary(neurorisk_life)
quantile(neurorisk_life,c(.025,.975))

summary(retinrisk_life_post)
quantile(retinrisk_life_post,c(.025,.975))
summary(retinrisk_life)
quantile(retinrisk_life,c(.025,.975))

summary(esrdrisk_life_post)
quantile(esrdrisk_life_post,c(.025,.975))
summary(esrdrisk_life)
quantile(esrdrisk_life,c(.025,.975))

summary(ascvd_life_post)
quantile(ascvd_life_post,c(.025,.975))
summary(ascvd_life)
quantile(ascvd_life,c(.025,.975))

summary(bca_life_post)
quantile(bca_life_post,c(.025,.975))
summary(bca_life)
quantile(bca_life,c(.025,.975))



```

# healthcare costs for employer and for medicare
Payer cost values:
MI, $4,372 private, $4,951 medicare (SE 525.6	977.5)
Stroke, $8,242 private,	$9,548 medicare (2,160.00	2,799.00)
Retinopathy, $801	private, $2,192 medicare (110.9	607.7)
Nephropathy, $5,224	private, $5,301 medicare (694.3	1,616.80)
Neuropathy, $1,996 private, $2,820 medicare (219.1	607.7)
Breast ca, $1,928 private, $1,612 medicare (326.7 184.6)

Member OOP costs:
MI: 232 (30.0)
Stroke, 446 (112.4)	
Retinopathy, 146 (20.3)
Nephropathy, 272 (44.2)
Neuropathy, 256 (25.3)
Breast ca, 226 (23.7)


```{r hlth cost}
emp_cost_mi = 4372
emp_cost_str = 8242
emp_cost_ascvd = mean(emp_cost_mi,emp_cost_str)
emp_cost_retin = 801
emp_cost_esrd = 5224
emp_cost_neuro = 1996
emp_cost_bca = 1928

mcare_cost_mi = 4951
mcare_cost_str = 9548
mcare_cost_ascvd = mean(mcare_cost_mi,mcare_cost_str)
mcare_cost_retin = 2192
mcare_cost_esrd = 5301
mcare_cost_neuro = 2820
mcare_cost_bca = 1612

disab = 0.14
turnover = 0.18

#emp portion of costs: 1-lifeexp_prop

emp_cost_neuro_tot = emp_cost_neuro *lifeexp* (neurorisk_life_post-neurorisk_life)*(1-lifeexp_prop)*(1/(1.03^lifeexp))*(1-disab)*(1-turnover)
mean(emp_cost_neuro_tot)
quantile(emp_cost_neuro_tot,c(.025,.975))

emp_cost_retin_tot = emp_cost_retin *lifeexp* (retinrisk_life_post-retinrisk_life)*(1-lifeexp_prop)*(1/(1.03^lifeexp))*(1-disab)*(1-turnover)
mean(emp_cost_retin_tot)
quantile(emp_cost_retin_tot,c(.025,.975))

emp_cost_esrd_tot = emp_cost_esrd *lifeexp* (esrdrisk_life_post-esrdrisk_life)*(1-lifeexp_prop)*(1/(1.03^lifeexp))*(1-disab)*(1-turnover)
mean(emp_cost_esrd_tot)
quantile(emp_cost_esrd_tot,c(.025,.975))

emp_cost_ascvd_tot = emp_cost_ascvd *lifeexp* (ascvd_life_post-ascvd_life)*(1-lifeexp_prop)*(1/(1.03^lifeexp))*(1-disab)*(1-turnover)
mean(emp_cost_ascvd_tot)
quantile(emp_cost_ascvd_tot,c(.025,.975))

emp_cost_bca_tot = emp_cost_bca *lifeexp* (bca_life_post-bca_life)*(1-lifeexp_prop)*(1/(1.03^lifeexp))*(1-disab)*(1-turnover)
mean(emp_cost_bca_tot)
quantile(emp_cost_bca_tot,c(.025,.975))


# adj for prev
prev_dm = 0.017
prev_ascvd = 0.055+0.074
prev_bca = 0.6254/100 #https://gis.cdc.gov/Cancer/USCS/DataViz.html

emp_cost_tot = prev_dm*(emp_cost_neuro_tot+emp_cost_retin_tot+emp_cost_esrd_tot)+
  prev_ascvd*emp_cost_ascvd_tot+
  prev_bca*emp_cost_bca_tot
summary(emp_cost_tot)
quantile(emp_cost_tot, c(.025,.975))

# Member portion of costs

mem_cost_mi = 232
mem_cost_str = 446
mem_cost_ascvd = mean(mem_cost_mi,mem_cost_str)
mem_cost_retin = 146
mem_cost_esrd = 272
mem_cost_neuro = 256
mem_cost_bca = 226


mem_cost_neuro_tot = mem_cost_neuro *lifeexp* (neurorisk_life_post-neurorisk_life)*(1/(1.03^lifeexp))
mean(mem_cost_neuro_tot)
quantile(mem_cost_neuro_tot,c(.025,.975))

mem_cost_retin_tot = mem_cost_retin *lifeexp* (retinrisk_life_post-retinrisk_life)*(1/(1.03^lifeexp))
mean(mem_cost_retin_tot)
quantile(mem_cost_retin_tot,c(.025,.975))

mem_cost_esrd_tot = mem_cost_esrd *lifeexp* (esrdrisk_life_post-esrdrisk_life)*(1/(1.03^lifeexp))
mean(mem_cost_esrd_tot)
quantile(mem_cost_esrd_tot,c(.025,.975))

mem_cost_ascvd_tot = mem_cost_ascvd *lifeexp* (ascvd_life_post-ascvd_life)*(1/(1.03^lifeexp))
mean(mem_cost_ascvd_tot)
quantile(mem_cost_ascvd_tot,c(.025,.975))

mem_cost_bca_tot = mem_cost_bca *lifeexp* (bca_life_post-bca_life)*(1/(1.03^lifeexp))
mean(mem_cost_bca_tot)
quantile(mem_cost_bca_tot,c(.025,.975))

mem_cost_tot = prev_dm*(mem_cost_neuro_tot+mem_cost_retin_tot+mem_cost_esrd_tot)+
  prev_ascvd*mem_cost_ascvd_tot+
  prev_bca*mem_cost_bca_tot
summary(mem_cost_tot)
quantile(mem_cost_tot, c(.025,.975))



#Medicare portion of costs: lifeexp_prop

mcare_cost_neuro_tot = mcare_cost_neuro *lifeexp* (neurorisk_life_post-neurorisk_life)*(1-lifeexp_prop)*(1+disab)*(1+turnover)
mean(mcare_cost_neuro_tot)
quantile(mcare_cost_neuro_tot,c(.025,.975))

mcare_cost_retin_tot = mcare_cost_retin *lifeexp* (retinrisk_life_post-retinrisk_life)*(1-lifeexp_prop)*(1+disab)*(1+turnover)
mean(mcare_cost_retin_tot)
quantile(mcare_cost_retin_tot,c(.025,.975))

mcare_cost_esrd_tot = mcare_cost_esrd *lifeexp* (esrdrisk_life_post-esrdrisk_life)*(1-lifeexp_prop)*(1+disab)*(1+turnover)
mean(mcare_cost_esrd_tot)
quantile(mcare_cost_esrd_tot,c(.025,.975))

mcare_cost_ascvd_tot = mcare_cost_ascvd *lifeexp* (ascvd_life_post-ascvd_life)*(1-lifeexp_prop)*(1+disab)*(1+turnover)
mean(mcare_cost_ascvd_tot)
quantile(mcare_cost_ascvd_tot,c(.025,.975))

mcare_cost_bca_tot = mcare_cost_bca *lifeexp* (bca_life_post-bca_life)*(1-lifeexp_prop)*(1+disab)*(1+turnover)
mean(mcare_cost_bca_tot)
quantile(mcare_cost_bca_tot,c(.025,.975))

mcare_cost_tot = prev_dm*(mcare_cost_neuro_tot+mcare_cost_retin_tot+mcare_cost_esrd_tot)+
  prev_ascvd*mcare_cost_ascvd_tot+
  prev_bca*mcare_cost_bca_tot
mean(mcare_cost_tot)
quantile(mcare_cost_tot, c(.025,.975))


```

# QALY

QALY disutility values:
appendix table 4 from: https://dx.doi.org/10.1016%2FS2213-8587(16)30270-4

Disability weights
MI, days 1-2 0.422 (0.284, 0.560)
MI, days 3-28 0.056 (0.035, 0.082)
Stroke, mild† 0.021 (0.011, 0.037)
Stroke, moderate 0.076 (0.050, 0.110)
Stroke, moderate with cognitive involvement 0.312 (0.211, 0.433)
Stroke, severe 0.539 (0.363, 0.705)
Stroke, severe with cognitive involvement 0.567 (0.394, 0.738)
Distance vision blindness 0.195 (0.132, 0.272)
Near vision impairment 0.013 (0.006, 0.024)
End-stage renal disease 0.573 (0.397, 0.749)
Diabetic neuropathy§ 0.099 (0.066, 0.145)
Amputation of one leg, long-term with treatment 0.021 (0.011, 0.035)
Amputation of one leg, long-term without treatment 0.164 (0.111, 0.229)
Breast cancer, Diagnosis and primary treatment, 0·288 (0·193–0·399)

```{r qalys}

mem_qaly_mi = (0.422*2+.056*28)/30
mem_qaly_str = (0.021*1/5+0.076*1/5+0.312*1/5+0.539*1/5+0.567*1/5)
mem_qaly_ascvd = mean(mem_qaly_mi,mem_qaly_str)
mem_qaly_retin = (0.195+0.013)/2
mem_qaly_esrd = 0.573
mem_qaly_neuro = (0.099/3+0.021/3+0.164/3)
mem_qaly_bca = 0.288


# calc method: https://academic.oup.com/heapol/article/21/5/402/578296
mem_qaly_neuro_tot = (1-mem_qaly_neuro)*(1-exp(-0.03*lifeexp/2))/0.03* (neurorisk_life_post-neurorisk_life)
mean(mem_qaly_neuro_tot)*1000
quantile(mem_qaly_neuro_tot,c(.025,.975))*1000

mem_qaly_retin_tot = (1-mem_qaly_retin) *(1-exp(-0.03*lifeexp/2))/0.03* (retinrisk_life_post-retinrisk_life)
mean(mem_qaly_retin_tot)*1000
quantile(mem_qaly_retin_tot,c(.025,.975))*1000

mem_qaly_esrd_tot = (1-mem_qaly_esrd) *(1-exp(-0.03*lifeexp/2))/0.03* (esrdrisk_life_post-esrdrisk_life)
mean(mem_qaly_esrd_tot)*1000
quantile(mem_qaly_esrd_tot,c(.025,.975))*1000

mem_qaly_ascvd_tot = (1-mem_qaly_ascvd) *(1-exp(-0.03*lifeexp/2))/0.03* (ascvd_life_post-ascvd_life)
mean(mem_qaly_ascvd_tot)*1000
quantile(mem_qaly_ascvd_tot,c(.025,.975))*1000

mem_qaly_bca_tot = (1-mem_qaly_bca) *(1-exp(-0.03*lifeexp/2))/0.03* (bca_life_post-bca_life)
mean(mem_qaly_bca_tot)*1000
quantile(mem_qaly_bca_tot,c(.025,.975))*1000

mem_qaly_tot = prev_dm*(mem_qaly_neuro_tot+mem_qaly_retin_tot+mem_qaly_esrd_tot)+
  prev_ascvd*mem_qaly_ascvd_tot+
  prev_bca*mem_qaly_bca_tot
mean(mem_qaly_tot)*1000
quantile(mem_qaly_tot, c(.025,.975))*1000


```
